---
title: "The best cities for buying a rental home"
format: 
  html:
    theme: lumen
    code-fold: true
    code-summary: "Show the code"
    include-in-header:
      - file: github-corner.html
execute:
  include: false
  messages: false
editor_options: 
  chunk_output_type: console
---

### Price to Income Ratio - The best cities for buying a rental home
Lower numbers are better for prospective homebuyers. For example, a \$500,000 home would rent for 4,200 (\$50k per year) at a 10 to 1 ratio, recuperating its value in 10 years. At 30 to 1, a \$500,000 home would rent at \$1,400.

```{r}
pacman::p_load(collapse, mapview)
load("data.rds")
```


```{r}

ratio_groups <- c(3, 10, 15, 20, 25, 30, Inf)
map_data <- census_data |> 
  fsubset(
    B25077_001E> 5e4 & !grepl("^((15)|([67])|(02))", GEOID)) |> 
  fmutate(
    `Price to Rent Ratio` = round(B25077_001E / (B25058_001E * 12), 0.1),
    Label = (cut(
      `Price to Rent Ratio`,
      include.lowest = TRUE,
      breaks = ratio_groups,
      labels = paste(ratio_groups[1:length(ratio_groups)-1], "to 1")
    )),
    across(c("B25077_001E", "B25058_001E"), scales::dollar)
  ) |>
  fsubset(!is.na(`Price to Rent Ratio`)) |> 
  # dplyr::slice_sample(n = 200) |>
  fselect(
    name = NAME, `Price to Rent Ratio`, `Home Value` = B25077_001E, 
    `Contract Rent` = B25058_001E, Label, geometry) |> 
  roworder(Label)

summary(map_data)
```

```{r}
#| include: true
#| echo: false
#| out-height: 80vh
#| out-width: 80vw
# hcl.pals("sequential")
pal <- hcl.colors(length(levels(map_data$Label)), "BuGn", rev = FALSE)
m <- mapview()
i <- 0
for (level in levels(map_data$Label)) {
  i <- i+1
  m <- m + mapview(
  map_data |> fsubset(Label == level),
  # zcol = "Label",
  # burst = TRUE,
  homebutton = FALSE,
  layer.name = level,
  col.regions = pal[i],
  color = colorspace::darken(pal[i], amount = 0.3),
  map.types = c(
    "CartoDB.Positron", "OpenStreetMap", "Esri.WorldImagery", "OpenTopoMap"
  ),
  lwd = 2
)
}
m
```

Source: 2022 5-year American Community Survey. Showing `r nrow(map_data) |> scales::comma()` places with home value and rent data, and median home values over $50k.